---
title: "Setting credentials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting credentials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we show you how to set up credentials for Ulysses. Each site has their own OMOP database and credentials. They should be stored securely. Likely you have the same credentials for every study, so you want to load them once. This vignette shows you how to set up the config.yml file automatically. 

```{r setup}
library(Ulysses)
```


## Make `shhh.csv`

First thing, we want to build our credentials table as a csv file. This should store your current database schemas, user and password to access the OMOP data. 


```{r}
#| label: initCred

initCredentials(
  db_id = "omop_dat",
  db_full_name = "My OMOP Database",
  dbms = "snowflake",
  user = "mike_jones",
  password = "who",
  connection_string = "jdbc:snowflake://host.com?remaining_string",
  cdm_database_schema = "cmd",
  work_database_schema = "scratch"
)

```

