---
title: {{{ Study }}} Analysis Plan 
number-sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Background

<!-------Add background on the study------>
# Research Questions

<!-------Include the study research questions------>


# Objectives


<!-------Add the study objective------>

## Primary Objectives

## Secondary Objectives

<!----Delete if not used------>

## Exploratory Objectives
<!----Delete if not used------>

# Study Design

## Study Population (Target Cohort)

<!---Provide information on the study population used in the analysis--->

## Exposure Definition

<!---Provide information on the exposure definition used in the analysis--->

## Outcome Definition

<!---Provide information on the outcome definition used in the analysis--->


# Data Sources

<!-- Add data sources, Example table for data sources -->


```{r}
#| label: tbl-data_sources
#| echo: false
#| tbl-cap: Data Sources used in study

dataSources <- tibble::tribble(
  ~Name, ~Country, ~Type, ~Size, ~Availability, ~Description,
  "Synpuf", "US", "claims", "110K", "2008-2010", "Synthetic data source based on US Medicare data"
)

flextable::flextable(dataSources)

```

# Analysis Plan

<!----Provide information on how the analysis will be done*---->


<!----Delete text that doesnt fit------->

## Cohort Diagnostics

Prior to running any specific analysis, we will evaluate the HMB cohort using the OHDSI R package `CohortDiagnostics`. This package produces metrics such as cohort counts in the database, incidence rates (by calendar year, age and gender), time distributions, cohort attrition and breakdown of index events. Evaluation of these metrics helps ensure that the clinical cohort is indeed reliable in capturing HMB in the OMOP database. 

## Baseline Characteristics


We assess baseline characteristics based on a observation window of 365 to 0 days prior to the index date. Categorical covariates are reported using the count and percentage. Continuous covariates are reported using the median, 25th and 75th percentile. Table xxx provides the set of covariates to be assessed in the study. 

**Demographics**

* Age as 5-year categories
* Age as continuous
* Race (if available)
* Ethnicity (if available)
* Year of Diagnosis (per calendar year)

<!----Add more sections for the analysis--------------------->

## Stratafication


<!----Add stratifications -------------->


# References

::: {#refs}
:::
